Spectroscopic confirmation of very-wide stellar binaries and large-separation comoving pairs from \textsl{Gaia DR1}

### 2017-08-14 paper/comoving-followup.tex
final edits

\author{Adrian~M.~Price-Whelan}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\author{Semyeong~Oh}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

@@

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique pairs and $\approx 120$
calibration targets.
All calibration targets have previously measured radial velocities from
high-resolution spectroscopic surveys (\citealt{Bensby:2014,Pinsonneault:2014})
in order to validate our radial velocity measurements.
These stars also have measured chemical abundances, and will be used in future
work as a training set for a method of transferring chemical abundance labels,
\project{The Cannon} (\citealt{Ness:2015}), to infer
metallicities and stellar parameters for our observed sources.
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python, described below.

\subsubsection{1D spectrum extraction}

\subsubsection{Wavelength calibration}

\subsubsection{Radial velocity measurement}

@@

The information contained in a large sample of widely-separated, comoving star
pairs spans from the dynamics of stars and dark matter in the Galaxy, the
fundamentals of star formation, and the stability of planetary systems.
Through our own low-resolution spectroscopic follow-up and a cross-match with
the RAVE \DR{5} catalog, we have identified \ncomovingtotal\ comoving star pairs
by combining these radial velocity measurements with astrometry from the \tgas\
catalog.
The separation distribution of these spectroscopically-confirmed comoving star
pairs presented shows an abundance of comoving pairs at separations between
$1$--$10~\pc$ (\figurename~\ref{fig:separation-RAVE}), the imposed limit in our
initial search (\citealt{Oh:2017}).
This result is further tentative evidence (see also \citealt{Shaya:2011}) for
the existence of a population of unbound comoving stars that represent the
ultimate fate of wide binaries and moving groups that slowly dissociate due to
weak perturbations from stars (\citealt{Jiang:2010}) and other massive
perturbers in the Galaxy.
With future data releases from the \gaia\ mission combined with large
spectroscopic surveys, the anticipated catalogs of confirmed comoving star pairs
will offer a new ``stereo'' view on the evolution and state of our Galaxy.

\todo{Reiterate that monotonically decreasing separation distribution is too
simple}

@@

It is a pleasure to thank
Ruth Angus (Columbia),
John Brewer (Yale),
Jason Curtis (Columbia),
Keith Hawkins (Columbia),
David W. Hogg (NYU),
Melissa Ness (MPIA), and
Dave Zurek (AMNH)
for useful discussions and advice.
We thank James Davenport (WWU) for releasing his open source spectroscopic
reduction code, \package{PyDIS}, which served as inspiration for the pipeline
used in this work.
This work is based on observations obtained at the MDM Observatory, operated by
Dartmouth College, Columbia University, Ohio State University, Ohio University,
and the University of Michigan.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.
This research has made use of the SIMBAD database, operated at CDS, Strasbourg,
France.
% The authors are pleased to acknowledge that the work reported on in this
% paper was substantially performed at the TIGRESS high performance computer
% center at Princeton University which is jointly supported by the Princeton
% Institute for Computational Science and Engineering and the Princeton
% University Office of Information Technology's Research Computing department.
Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic
Telescope LAMOST) is a National Major Scientific Project built by the Chinese
Academy of Sciences. Funding for the project has been provided by the National
Development and Reform Commission. LAMOST is operated and managed by the
National Astronomical Observatories, Chinese Academy of Sciences.

@@

>>>>

\author[0000-0003-0872-7098]{Adrian~M.~Price-Whelan}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

\author[0000-0001-7790-5308]{Semyeong~Oh}
\affiliation{Department of Astrophysical Sciences,
             Princeton University, Princeton, NJ 08544, USA}

@@

We observed a total of 765 unique sources over 5 nights from (UT) 2017-03-11 to
2017-03-15, corresponding to 323 unique pairs and $\approx 120$
calibration targets.
All calibration targets have previously measured radial velocities from
high-resolution spectroscopic surveys (\citealt{Bensby:2014,Pinsonneault:2014})
in order to validate our radial velocity measurements.
These stars also have measured chemical abundances, and will be used in future
work as a training set for a method of transferring chemical abundance labels,
\project{The Cannon} (\citealt{Ness:2015}), to infer
metallicities and stellar parameters for our observed sources.
The spectra were reduced and calibrated using a custom, publicly-available
pipeline written in \python; see Appendices~\ref{sec:1d-ext}--\ref{sec:rv-meas}
for a full description of the data reduction procedure.

@@

The information contained in a large sample of widely-separated, comoving star
pairs spans from the dynamics of stars and dark matter in the Galaxy, the
fundamentals of star formation, and the stability of planetary systems.
Through our own low-resolution spectroscopic follow-up and a cross-match with
the RAVE \DR{5} catalog, we have identified \ncomovingtotal\ comoving star pairs
by combining these radial velocity measurements with astrometry from the \tgas\
catalog.
From a cross-match to the \acronym{LAMOST} survey, we find that candidate
comoving star pairs with small differences in radial velocities also have small
differences in iron abundance, even at large separations ($>1~\pc$).
The separation distribution of the spectroscopically-confirmed comoving star
pairs presented shows an abundance of comoving pairs at separations between
$1$--$10~\pc$ (\figurename~\ref{fig:separation-RAVE}), the imposed limit in our
initial search (\citealt{Oh:2017}).
This result is further tentative evidence (see also \citealt{Shaya:2011}) for
the existence of a population of unbound comoving stars that represent the
ultimate fate of wide binaries and moving groups that slowly dissociate due to
weak perturbations from stars (\citealt{Jiang:2010}) and other massive
perturbers in the Galaxy.
From this work it appears that modeling the population of wide binaries or
comoving stars with a monotonically decreasing separation distribution---as is commonly done---too simple to also represent the population of unbound systems.
With future data releases from the \gaia\ mission combined with large
spectroscopic surveys, the anticipated catalogs of confirmed comoving star pairs
will offer a new ``stereo'' view on the evolution and state of our Galaxy.

@@

It is a pleasure to thank
Ruth Angus (Columbia),
John Brewer (Yale),
Jason Curtis (Columbia),
Keith Hawkins (Columbia),
David W. Hogg (NYU),
Melissa Ness (MPIA), and
Dave Zurek (AMNH)
for useful discussions and advice.
We thank James Davenport (WWU) for releasing his open source spectroscopic
reduction code, \package{PyDIS}, which served as inspiration for the pipeline
used in this work.
This work is based on observations obtained at the MDM Observatory, operated by
Dartmouth College, Columbia University, Ohio State University, Ohio University,
and the University of Michigan.
This work has made use of data from the European Space Agency (ESA)
mission {\it Gaia} (\url{https://www.cosmos.esa.int/gaia}), processed by
the {\it Gaia} Data Processing and Analysis Consortium (DPAC,
\url{https://www.cosmos.esa.int/web/gaia/dpac/consortium}). Funding
for the DPAC has been provided by national institutions, in particular
the institutions participating in the {\it Gaia} Multilateral Agreement.
This research has made use of the SIMBAD database, operated at CDS, Strasbourg,
France.
This project was developed in part at the 2017 Heidelberg Gaia Sprint, hosted by the Max-Planck-Institut für Astronomie, Heidelberg.

Guoshoujing Telescope (the Large Sky Area Multi-Object Fiber Spectroscopic
Telescope LAMOST) is a National Major Scientific Project built by the Chinese
Academy of Sciences. Funding for the project has been provided by the National
Development and Reform Commission. LAMOST is operated and managed by the
National Astronomical Observatories, Chinese Academy of Sciences.

@@

\appendix

\section{1D spectrum extraction} \label{sec:1d-ext}

\section{Wavelength calibration}

\section{Radial velocity measurement} \label{sec:rv-meas}
